Four decades of water quality change in the upper San Francisco Estuary\footnote{Version: Thu Feb 9 09:38:49 2017 -0600, \href{https://github.com/fawda123/sftrends_manu/commit/dc55a82ecaa6c7579c80e824377410a6b61dc53f}{dc55a82ecaa6c7579c80e824377410a6b61dc53f}}

### 2017-05-02 sftrends_manu.Rnw
fixed phil's name

\author{Marcus W. Beck}
\affiliation{USEPA National Health and Environmental Effects Research Laboratory, Gulf Ecology Division, Gulf Breeze, FL}
\email{beck.marcus@epa.gov}
\phone{+1 (850)9342480}
\fax{+1 (850)9342401}
\author{David B. Senn}
\author{Thomas W. Jabusch}
\author{Phil Trowbridge}
\affiliation{San Francisco Estuary Institute, Richmond, CA}

@@

Relationships of nirite/nitrate with flow described by \ac{wrtds} showed an inverse flow and concentration dynamic with flushing or dilution at higher flow (\cref{fig:p8trnds}b).  Seasonal variation was even more apparent for ammonium, although both nitrite/nitrate and ammonium typically had the highest concentrations at low flow in the winter (January).  Additionally, strength of the flow/nutrient relationship changed between years. Nitrite/nitrate typically had the strongest relationship with flow later in the time series (i.e., larger negative slope), whereas ammonium had the strongest relationship with flow around 2000 in January.  Using \ac{wrtds}, an empirical link is created between upstream changes and observed effects downstream that is characterized by differences in analytes between years and season. A general conclusion is that ammonium reductions were concurrent with \ac{wwtp} upgrades, but the reduction was most apparent at low-flow in January.  These dynamics are difficult to characterize from the observed time series, and further, results from \ac{wrtds} can be used to develop additional hypotheses of factors that influence nutrient concentrations at P8. For example, estimated ammonium concentrations in July were low for all flow levels which suggests either nitrogen inputs were low in the summer or nitrogen was available and uptake by primary consumers was high. Seasonal patterns in the relationship between flow and nitrite/nitrate were not as dramatic as compared to ammonium, and in particular, low-flow events in July were generally associated with higher concentrations.  This could suggest that ammonium concentrations at P8 are driving phytoplankton production at low flow during warmer months, and not nitrite/nitrate given the higher estimated concentrations in July at low flow. As such, these simple observations from \ac{wrtds} provide quantitative support of cause and effect mechanisms of nutrient impacts on potentially adverse environmental conditions as they relate to nutrient-related source controls upstream.

>>>>

\author{Marcus W. Beck}
\affiliation{USEPA National Health and Environmental Effects Research Laboratory, Gulf Ecology Division, Gulf Breeze, FL}
\email{beck.marcus@epa.gov}
\phone{+1 (850)9342480}
\fax{+1 (850)9342401}
\author{David B. Senn}
\author{Thomas W. Jabusch}
\author{Philip R. Trowbridge}
\affiliation{San Francisco Estuary Institute, Richmond, CA}

@@

Relationships of nitrite/nitrate with flow described by \ac{wrtds} showed an inverse flow and concentration dynamic with flushing or dilution at higher flow (\cref{fig:p8trnds}b).  Seasonal variation was even more apparent for ammonium, although both nitrite/nitrate and ammonium typically had the highest concentrations at low flow in the winter (January).  Additionally, strength of the flow/nutrient relationship changed between years. Nitrite/nitrate typically had the strongest relationship with flow later in the time series (i.e., larger negative slope), whereas ammonium had the strongest relationship with flow around 2000 in January.  Using \ac{wrtds}, an empirical link is created between upstream changes and observed effects downstream that is characterized by differences in analytes between years and season. A general conclusion is that ammonium reductions were concurrent with \ac{wwtp} upgrades, but the reduction was most apparent at low-flow in January.  These dynamics are difficult to characterize from the observed time series, and further, results from \ac{wrtds} can be used to develop additional hypotheses of factors that influence nutrient concentrations at P8. For example, estimated ammonium concentrations in July were low for all flow levels which suggests either nitrogen inputs were low in the summer or nitrogen was available and uptake by primary consumers was high. Seasonal patterns in the relationship between flow and nitrite/nitrate were not as dramatic as compared to ammonium, and in particular, low-flow events in July were generally associated with higher concentrations.  This could suggest that ammonium concentrations at P8 are driving phytoplankton production at low flow during warmer months, and not nitrite/nitrate given the higher estimated concentrations in July at low flow. As such, these simple observations from \ac{wrtds} provide quantitative support of cause and effect mechanisms of nutrient impacts on potentially adverse environmental conditions as they relate to nutrient-related source controls upstream.

### 2017-05-02 sftrends_manu.tex
fixed phil's name

\author{Marcus W. Beck}
\affiliation{USEPA National Health and Environmental Effects Research Laboratory, Gulf Ecology Division, Gulf Breeze, FL}
\email{beck.marcus@epa.gov}
\phone{+1 (850)9342480}
\fax{+1 (850)9342401}
\author{David B. Senn}
\author{Thomas W. Jabusch}
\author{Phil Trowbridge}
\affiliation{San Francisco Estuary Institute, Richmond, CA}

\title[Water quality change in the upper SFE]{Four decades of water quality change in the upper San Francisco Estuary\footnote{Version: Tue Mar 14 09:11:12 2017 -0500, \href{https://github.com/fawda123/sftrends_manu/commit/7b5a17c006b01d83e6f284649e548d329f7ada14}{7b5a17c006b01d83e6f284649e548d329f7ada14}\newline The views expressed in this paper are those of the authors and do not necessarily reflect the views or policies of the U.S. Environmental Protection Agency.}}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}
\linenumbers

>>>>

\author{Marcus W. Beck}
\affiliation{USEPA National Health and Environmental Effects Research Laboratory, Gulf Ecology Division, Gulf Breeze, FL}
\email{beck.marcus@epa.gov}
\phone{+1 (850)9342480}
\fax{+1 (850)9342401}
\author{David B. Senn}
\author{Thomas W. Jabusch}
\author{Philip R. Trowbridge}
\affiliation{San Francisco Estuary Institute, Richmond, CA}

\title[Water quality change in the upper SFE]{Four decades of water quality change in the upper San Francisco Estuary\footnote{Version: Wed Mar 15 17:49:38 2017 -0500, \href{https://github.com/fawda123/sftrends_manu/commit/835c0b4e439bcdbb0228b9b56799bfb4764b7d97}{835c0b4e439bcdbb0228b9b56799bfb4764b7d97}\newline The views expressed in this paper are those of the authors and do not necessarily reflect the views or policies of the U.S. Environmental Protection Agency.}}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}
\linenumbers
